---
title: "Checking Assumption Violations"
author: "Drusilla Talawa"
date: "9/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulating Some Data

```{r}
set.seed(6832)

normal.small <- rnorm(15, mean=5, sd=2)
normal.large <- rnorm(250, mean=5, sd=2)

uniform <- runif(50, min=10, max=20)
chisq <- rchisq(50, df=1.5)

outlier <- c(rpois(49, lambda=10),50)

myData <- data.frame(Group = factor(rep(LETTERS[1:3], each=30)),
                     Outcome = c(rnorm(30, mean=2, sd=.4),
                                 rpois(30, lambda=2),
                                 runif(30, min=1, max=6)))

myData.ub <- myData[-sample(nrow(myData),size=10),]

myData.same <- data.frame(Group = factor(rep(LETTERS[1:3], each=30)),
                     Outcome = c(rnorm(30, mean=4, sd=1.414),
                                 rpois(30, lambda=2),
                                 runif(30, min=1, max=6)))

myData.same.ub <- data.frame(Group = factor(c(rep("A",8),rep("B",25),rep("C",20))),
                     Outcome = c(rnorm(8, mean=4, sd=1.414),
                                 rpois(25, lambda=2),
                                 runif(20, min=1, max=6)))
```

\newpage
## Normality

### A True Case: Actually Normal, Large n

``` {r}
#this won't work for ggplot, it's a way to organize data into a grid
par(mfrow=c(1, 2)) ##c(row, column)
## Histograms + Normal Curve
hist(normal.large, freq = FALSE) ##because hist is probability = 1)
lines(density(normal.large))
curve(dnorm(x, mean=mean(normal.large), sd=sd(normal.large)), col="red", lwd=2, add=TRUE)

##QQ plots
qqnorm(normal.large)
qqline(normal.large) #for the one-to-one line
#this is considered a good QQ plot, very normal

##Skew & Kurtosis
##install.packages("pastecs")
##library(pastecs)
stat.desc(x=normal.large, norm = TRUE)

#z_skew = skew/SE compared to critical values in N(0, 1)
#z*_skew = skew/2*SE(standard error), compare this to 1

shapiro.test(x=normal.large)
#p.value = 0.70 > any reasonable alpha, fail to reject Ho. 
#Conclude that the data may be approximately normal.

```


### A True Small Case: Actually Normal, Small n

``` {r}

par(mfrow=c(1, 2))

hist(normal.small, freq = FALSE)

lines(density(normal.small))

curve(dnorm(x, mean=mean(normal.small), sd=sd(normal.small)), col="red", lwd=2, add=TRUE)

##QQ plots
qqnorm(normal.small)
qqline(normal.small) 

stat.desc(x=normal.small, norm = TRUE)

shapiro.test(x=normal.small)
#p-value = 0.41 > any reasonable alpha, fail to reject Ho. 
#conclude that the data may be appx. normal.

#We cannot prove normality, we can say appx. or roughly
#But when we do reject Ho, we can say that it's not normal. Because in this test we are assuming the distribution is normal. 
```

### False Cases: When It's Not Normal

``` {r}

par(mfrow=c(1, 2))

hist(chisq, freq = FALSE)

lines(density(chisq))
curve(dnorm(x, mean=mean(chisq), sd=sd(chisq)), col="red", lwd=2, add=TRUE)

##QQ plots
#trying to get data that are 4 sd above and 4 sd below. to make all the data appear in the graphic
qqnorm(chisq, xlim=c(-4, 4), ylim=mean(chisq)+c(-4,4)*sd(chisq))
qqline(chisq) 

round(stat.desc(x=chisq, norm = TRUE), digits = 4)
#round in any function that produces numerical output, to change the scientific notation to normal numbers 

(shapiro.test(x=chisq))
#p-value is appx. 0 which is < any reasonable alpha, REJECT Ho. 
#conclude that the data is NOT Normal.
```


First, we'll see what happens with asymmetric data.

Next we will see what happens with a flat distribution.

``` {r}

par(mfrow=c(1, 2))

hist(uniform, freq = FALSE)

lines(density(uniform))
curve(dnorm(x, mean=mean(uniform), sd=sd(uniform)), col="red", lwd=2, add=TRUE)

##QQ plots
#trying to get data that are 4 sd above and 4 sd below. to make all the data appear in the graphic
qqnorm(uniform, xlim=c(-4, 4), ylim=mean(uniform)+c(-4,4)*sd(uniform))
qqline(uniform) 

round(stat.desc(x=uniform, norm = TRUE), digits = 4)
#round in any function that produces numerical output, to change the scientific notation to normal numbers 

(shapiro.test(x=uniform))
#p-value is .0025 < most appropriate alpha, REJECT Ho. 
#conclude that the data is NOT Normal.
```

\newpage
## Homogeneity

### Balanced Groups with Equal Variance


### Balanced Groups with Unequal Variance


### Unbalanced Groups with Equal Variance


### Unbalanced Groups with Unequal Variance



\newpage
## Adjusting the Data

### Sensitivity Analyses

### Transforming a Variable